A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study

### 2022-08-05 02_paper/02_study/02_fitbit paper/fitbit_paper.Rmd
fitbit paper participants

author: 
  - name          : "Mandy Klatt"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Dittrichring 5-7, 04109 Leipzig"
    email         : "mandy.klatt@uni-leipzig.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Peer Keßler"
    affiliation   : "1"
    role:
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Gregor Kachel"
    affiliation   : "1,2"
    role:
      - "Conceptualization"
      - "Supervision"

@@

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

@@

@@

## Participants

>>>>

author: 
  - name          : "Mandy Klatt"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Dittrichring 5-7, 04109 Leipzig"
    email         : "mandy.klatt@uni-leipzig.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Peer Keßler"
    affiliation   : "1"
    role:
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Gregor Kachel"
    affiliation   : "1,2"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
  - name          : "Anne Deiglmayr"
    affiliation   : "1"
    role:
      - "Supervision"

@@

authornote: |
  We received funding from QualiFond of University Leipzig. We have no conflicts of interest to disclose. This article is based on data used at conference presentations (DACH-Nachwuchsakademie, 2022; EARLI SIG 11, 2022; EARLI SIG 27, 2022).

@@

```{r include=FALSE}
# install.packages("needs")

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

library(needs)

needs(tidyverse,
      lubridate,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable)

```

@@

## Participants
```{r demographicstable, include = TRUE, echo=FALSE}

# read in data and select and rename relevant columns with transmute 
demo.data <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% 
  transmute(Group = LI06_05,
            Gender = LI02_01_1,
            Age = LI03_01,
            `Teaching Experience` = LI04_01,
            School_type = LI18,
            Grade = LI16_05,
            Subject = LI10_05,
            Semester = LI11_01,
            Internship = LI13_01,
            `Extracurricular Teaching Experience` = LI14,
            `Secondary Teaching Activities` = LI17
            )

# define expert and novice with ifelse function
demo.data$Group = ifelse(demo.data$Group < 200, "Novice","Expert")

# changing gender to count female gender (male = 2 --> 0)
demo.data$Gender = ifelse(demo.data$Gender == 2, "1","0")

# changing character Gender into numeric
demo.data$Gender <- as.numeric(demo.data$Gender)

# creating new column with total participants to calculate percent of gender
demo.data <- demo.data %>%
  group_by(Group) %>%
  mutate(Sum_participants = n())

# replacing 2 to 1 (Oberschule = Gymnasisium --> Secondary school)
demo.data$School_type[demo.data$School_type == 2] <- "1"

# creating new column to calculate percent of school type
demo.data <- 
  demo.data %>%
  group_by(School_type, Group) %>%
  mutate(Count_Schooltype = sum(n())) %>% 
  mutate(Percent_schooltype = round((Count_Schooltype/Sum_participants) * 100, 
         digits = 2))

# changing character Percent_schooltype into numeric
demo.data$Percent_schooltype <- as.numeric(demo.data$Percent_schooltype)

# changing character Semester into numeric
demo.data$Semester <- as.numeric(demo.data$Semester)

# changing Type of School into character
# LI18:
# 1: Grundschule
# 2: Oberschule
# 3: Gymnasium
# 4: Förderschule/Sondershausen
# 5: Berufsbildende Schule
# -1: andere Schulform
demo.data$School_type <- factor(x = demo.data$School_type,
                   levels = c("1","3","4","5","-1"), # Mögliche Werte # andere Möglichkeit: unique(demo.data$School_type)
                   labels = c("Primary School","Secondary School", "Special Education","Vocational School", "Other Schooltype")) # Bezeichnungen

# changing `Extracurricular Teaching Experience` to count experience (no experience = 2 --> 0)
demo.data$`Extracurricular Teaching Experience` = ifelse(demo.data$`Extracurricular Teaching Experience` == 1, "1","0")

# changing character Extra. Exp. into numeric
demo.data$`Extracurricular Teaching Experience` <- as.numeric(demo.data$`Extracurricular Teaching Experience`)

# changing `Secondary Teaching Activities` to count experience (no experience = 2 --> 0)
demo.data$`Secondary Teaching Activities` = ifelse(demo.data$`Secondary Teaching Activities` == 1, "1","0")

# changing character Extra. Exp. into numeric
demo.data$`Secondary Teaching Activities` <- as.numeric(demo.data$`Secondary Teaching Activities`)

# create new columns that give proportion of Gender
# percent = female gender divided by all female Gender
# demo.data <- demo.data  %>% 
#   group_by(Group) %>% 
#   mutate(Gender_percent = Gender/Sum_participants) %>% 
#   mutate(Gender_percent = round(sum(Gender_percent)*100, digits=0))

# create a basic table (tibble) using tidyverse functions
demo.table <- demo.data %>%
  group_by(Group) %>%
  summarise(N=n(),
            "Gender female in percent" = round(sum(Gender/Sum_participants) * 100, 
                                               digits = 2),
            "M Age in years" = round(mean(Age), 
                                     digits = 2),
            "SD Age in years" = round(sd(Age), 
                                      digits = 2),
            "Min Age in years" = min(Age),
            "Max Age in years" = max(Age)
            )

# # format and insert table in manuscript
# demo.table %>%
# papaja::apa_table(
#   caption = "Demographic Information",
#   # note = "Write Note here",
#   escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
#   placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
#   font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
#   )

# create a basic table (tibble) using tidyverse functions
exp.table <- demo.data %>%
  group_by(Group) %>%
  summarise(N=n(),
           "M Exp." = round(mean(`Teaching Experience`), 2),
            "SD Exp." = round(sd(`Teaching Experience`), 2),
            "Min Exp." = min(`Teaching Experience`),
            "Max Exp." = max(`Teaching Experience`),
            "M Semester" = round(mean(Semester), 2),
            "SD Semester" = round(sd(Semester), 2),
            "Min Semester" = min(Semester),
            "Max Semester" = max(Semester),
            "M Internship" = round(mean(Internship), 2),
            "SD Internship" = round(sd(Internship), 2),
            "Min Internship" = min(Internship),
            "Max Internship" = max(Internship),
            "Extracurricular Teaching Experience" = round(sum(`Extracurricular Teaching Experience`/Sum_participants) * 100, digits = 2),
            "Secondary Teaching Activities" = round(sum(`Secondary Teaching Activities`/Sum_participants) * 100)
           )

# # format and insert table in manuscript
# exp.table %>%
# papaja::apa_table(
#   caption = "Teaching Experience in years, internship experience in teaching units (45min) and extracurricular teaching experience in percent",
#   # note = "Write Note here",
#   escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
#   placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
#   font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
#   )

```

XXX  fitbit data exclusion XXX

The sample consisted of *N* = `r length(demo.data$Group)` participants with *n* = `r length(demo.data$Group[demo.data$Group == "Expert"])` expert teachers and *n* = `r length(demo.data$Group[demo.data$Group == "Novice"])` novice teachers.

The inclusion criterion for experts was that they have successfully completed teacher training and are actively employed in the teaching profession. According to @palmer2005identifying, we selected teachers as experts who had at least three years of professional experience and ideally had worked in another teaching position, such as subject advisor or trainer for trainee teachers, in addition to their teaching profession in school. Novices were student teachers who had successfully completed their first internship in a school and gained one to four hours of teaching experience.

The expert teachers (`r demo.data %>% filter(Gender == 1 & Group == "Expert") %>% pull(Gender) %>% length()` women; `r demo.table %>% filter(Group == "Expert") %>% pull("Gender female in percent")`%) had a mean age of `r demo.data %>% filter(Group == "Expert") %>% pull(Age) %>% mean() %>% round(., digits = 1)` years (*SD* = `r demo.table %>% filter(Group == "Expert") %>% pull("SD Age in years") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Expert") %>% pull("Min Age in years")`-`r demo.table %>%  filter(Group == "Expert") %>% pull("Max Age in years")`) and an average teaching experience of `r exp.table %>% filter(Group == "Expert") %>% pull("M Exp.") %>% mean() %>% round(., digits = 1)` years (*SD* = `r exp.table %>% filter(Group == "Expert") %>% pull("SD Exp.") %>% round(., digits = 1)`; range: `r exp.table %>%  filter(Group == "Expert") %>% pull("Min Exp.")`-`r exp.table %>%  filter(Group == "Expert") %>% pull("Max Exp.")`). 
`r demo.data %>% filter(Group == "Expert" & School_type == "Primary School") %>% pull("Percent_schooltype") %>% unique()`% of the experts were primary school teachers and `r demo.data %>% filter(Group == "Expert" & School_type == "Secondary School") %>% pull("Percent_schooltype") %>% unique()`% were secondary school teachers. `r exp.table %>% filter(Group == "Expert") %>% pull("Secondary Teaching Activities")`% of the experienced teachers were also engaged in an secondary teaching activity, such as lecturers at the university, main training supervisors for trainee teachers and subject advisers.

The novice teachers (`r demo.data %>% filter(Gender == 1 & Group == "Novice") %>% pull(Gender) %>% length()` women; `r demo.table %>% filter(Group == "Novice") %>% pull("Gender female in percent")`%) had a mean age of `r demo.data %>% filter(Group == "Novice") %>% pull(Age) %>% mean() %>% round(., digits = 1)` years (*SD* = `r demo.table %>% filter(Group == "Novice") %>% pull("SD Age in years") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Novice") %>% pull("Min Age in years")`-`r demo.table %>%  filter(Group == "Novice") %>% pull("Max Age in years")`) with an average teaching experience of `r exp.table %>% filter(Group == "Novice") %>% pull("M Exp.") %>% mean() %>% round(., digits = 1)` years. On average, the student teachers were in their `r exp.table %>% filter(Group == "Novice") %>% pull("M Semester") %>% mean() %>% round(., digits = 1)` semester (*SD* = `r exp.table %>% filter(Group == "Novice") %>% pull("SD Semester") %>% round(., digits = 1)`; range: `r exp.table %>%  filter(Group == "Novice") %>% pull("Min Semester")`-`r exp.table %>%  filter(Group == "Novice") %>% pull("Max Semester")`). Furthermore, they had an average teaching experience of `r exp.table %>% filter(Group == "Novice") %>% pull("M Internship") %>% mean() %>% round(., digits = 1)` teaching units à 45min (*SD* = `r exp.table %>% filter(Group == "Novice") %>% pull("SD Internship") %>% round(., digits = 1)`; range: `r exp.table %>%  filter(Group == "Novice") %>% pull("Min Internship")`-`r exp.table %>%  filter(Group == "Novice") %>% pull("Max Internship")`) through the internships during their studies. 
`r demo.data %>% filter(Group == "Novice" & School_type == "Primary School") %>% pull("Percent_schooltype") %>% unique()`% of the novices were studying to become primary school teachers, `r demo.data %>% filter(Group == "Novice" & School_type == "Secondary School") %>% pull("Percent_schooltype") %>% unique()`% to become secondary school teachers and `r demo.data %>% filter(Group == "Novice" & School_type == "Special Education") %>% pull("Percent_schooltype") %>% unique()`% to become special education teachers. `r exp.table %>% filter(Group == "Novice") %>% pull("Extracurricular Teaching Experience")`% of the student teachers were also engaged in an extracurricular teaching activity, such as tutoring or homework supervision.

The subjects were primarily recruited through personal contacts, social media (Facebook), e-mail distribution lists and advertising in lectures at Leipzig University. All study procedures were carried out in accordance with the ethical standards of the University’s Institutional Review Board. The authors received a positive vote on the study procedures from the Ethics Committee Board of Leipzig University. All participants were informed in detail about the aim and intention of the study prior to testing. Participation in the study was voluntary and only took place after written consent has been given.

### 2022-08-05 02_paper/02_study/02_fitbit paper/fitbit_paper.tex
fitbit paper participants

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study},
  pdfauthor={Mandy Klatt1, Peer Keßler1, \& Gregor Kachel1,2},
  pdflang={en-EN},
  pdfkeywords={heart rate; photoplethysmography; wearable electronic device; expertise differences},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study}
\author{Mandy Klatt\textsuperscript{1}, Peer Keßler\textsuperscript{1}, \& Gregor Kachel\textsuperscript{1,2}}
\date{}

@@

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. Mandy Klatt: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Peer Keßler: Writing - Review \& Editing, Supervision; Gregor Kachel: Supervision.

@@

@@

>>>>

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study},
  pdfauthor={Mandy Klatt1, Peer Keßler1, Gregor Kachel1,2, \& Anne Deiglmayr1},
  pdflang={en-EN},
  pdfkeywords={heart rate; photoplethysmography; wearable electronic device; expertise differences},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study}
\author{Mandy Klatt\textsuperscript{1}, Peer Keßler\textsuperscript{1}, Gregor Kachel\textsuperscript{1,2}, \& Anne Deiglmayr\textsuperscript{1}}
\date{}

@@

We received funding from QualiFond of University Leipzig. We have no conflicts of interest to disclose. This article is based on data used at conference presentations (DACH-Nachwuchsakademie, 2022; EARLI SIG 11, 2022; EARLI SIG 27, 2022).

The authors made the following contributions. Mandy Klatt: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Peer Keßler: Writing - Original Draft Preparation, Writing - Review \& Editing; Gregor Kachel: Conceptualization; Anne Deiglmayr: Supervision.

@@

The sample consisted of \emph{N} = 41 participants with \emph{n} = 16 expert teachers and \emph{n} = 25 novice teachers.

The inclusion criterion for experts was that they have successfully completed teacher training and are actively employed in the teaching profession. According to Palmer, Stough, Burdenski, and Gonzales (2005), we selected teachers as experts who had at least three years of professional experience and ideally had worked in another teaching position, such as subject advisor or trainer for trainee teachers, in addition to their teaching profession in school. Novices were student teachers who had successfully completed their first internship in a school and gained one to four hours of teaching experience.

The expert teachers (9 women; 56.25\%) had a mean age of 41.40 years (\emph{SD} = 10.70; range: 27-59) and an average teaching experience of 13.60 years (\emph{SD} = 12.80; range: 2-37).
18.75\% of the experts were primary school teachers and 81.25\% were secondary school teachers. 56\% of the experienced teachers were also engaged in an secondary teaching activity, such as lecturers at the university, main training supervisors for trainee teachers and subject advisers.

The novice teachers (16 women; 64\%) had a mean age of 23.40 years (\emph{SD} = 1.80; range: 20-27) with an average teaching experience of 0 years. On average, the student teachers were in their 7.30 semester (\emph{SD} = 2.60; range: 3-11). Furthermore, they had an average teaching experience of 11.20 teaching units à 45min (\emph{SD} = 8.10; range: 0-36) through the internships during their studies.
20\% of the novices were studying to become primary school teachers, 68\% to become secondary school teachers and 12\% to become special education teachers. 88\% of the student teachers were also engaged in an extracurricular teaching activity, such as tutoring or homework supervision.

The subjects were primarily recruited through personal contacts, social media (Facebook), e-mail distribution lists and advertising in lectures at Leipzig University. All study procedures were carried out in accordance with the ethical standards of the University's Institutional Review Board. The authors received a positive vote on the study procedures from the Ethics Committee Board of Leipzig University. All participants were informed in detail about the aim and intention of the study prior to testing. Participation in the study was voluntary and only took place after written consent has been given.

@@

\leavevmode\vadjust pre{\hypertarget{ref-palmer2005identifying}{}}%
Palmer, D. J., Stough, L. M., Burdenski, T. K., Jr, \& Gonzales, M. (2005). Identifying teacher expertise: An examination of researchers' decision making. \emph{Educational Psychologist}, \emph{40}(1), 13--25.
